Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-07-31 Manuscript/MPESubmission/Revision/response.tex

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\pagenumbering{arabic}

%Pagination style and stuff
\linespread{2}
\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\begin{document}

Dear Dr D\'{a}valos,

MPE-15-254: Interim Decision

We are very grateful to both referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper. We have taken all of their comments on board, and respond to their points below. For improving clarity in this document, we sorted the reviewers comments as follows:
\begin{enumerate}
\item \textbf{Reviewer 1:} containing reviewer's 1 major comments
\item \textbf{Reviewer 1 Specific comments:} containing reviewer's 1 specific comments
\item \textbf{Reviewer 2:} containing reviewer's 2 comments
\end{enumerate}
Also, we uploaded two versions of the revised manuscript. Both have the exact same content but one has all the changes and additions to the main text highlighted in yellow.

% TG: actually, is it possible to send them the response as a pdf or will I have to copy paste all that in a mail? If the second, it's not that a big problem (just annoying) but I think we should try sending the pdf, it's way clearer like that.

@@

\section{Reviewer 1:}

\begin{enumerate}
\item{\textcolor{blue}{\textbf{Novelty:} The novelty of this study should be spelled out more clearly by more a more thorough review of the literature, not just by referring to a few relevant works, but by pointing out what is so new about this study.
The authors assert that this is the first study to examine the effects of missing morphological data from a total evidence dataset, but that is what Pattinson et al. 2014 did (as well as Wiens 2005).
This study examines it in a different way with a study design that is unique and adds to the field. For example, one feature of this study that is unique is that a tree with fossils is simulated and the data simulated on that tree.
In other studies of missing data, data were removed from living taxa to determine if their positions could still be inferred (Pattinson et al. 2015; Wiens and Tiu 2012).
This study will have bearing on an old assertion that including fossils specifically may help alleviate long-branch attraction, not only additional taxa with varying amounts of missing data, as suggested by others (e.g., Wiens 2005).
Similarly, in the discussion, the authors need to point out how their results fit in with the results of previous research more clearly. }}

% The novelty of this study should be spelled out more clearly by more a more thorough review of the literature, not just by referring to a few relevant works, but by pointing out what is so new about this study.

% The authors assert that this is the first study to examine the effects of missing morphological data from a total evidence dataset, but that is what Pattinson et al. 2014 did (as well as Wiens 2005).

% This study will have bearing on an old assertion that including fossils specifically may help alleviate long-branch attraction, not only additional taxa with varying amounts of missing data, as suggested by others (e.g., Wiens 2005).

% Similarly, in the discussion, the authors need to point out how their results fit in with the results of previous research more clearly.

@@

We gladly followed reviewer 1's graphical abstract suggestions and have changed it to be similar to the figure 5. % TG: It's freaking not catchy at all but at this point I give up. I've tried many ways but the flippin' paper IS NOT attractive. It might be useful, well written, interesting, citable, etc... But it won't be sexy.

\item{\textcolor{blue}{\textbf{Highlights:} wording seems awkward}}

We rephrased the highlights of the paper as follows: %3 to 5 bullet points (maximum 85 characters, including spaces, per bullet point). TG: probably need improvement here.

\begin{enumerate}[(1)]
\item The Total Evidence methods allows to combine living and fossil data.
\item This study test the effect of missing morphological data on topology.
\item The number of living taxa with data as the most effect on recovering topology.
\item Bayesian inference outperforms Maximum Likelihood for correct clade recovery.
\item We suggest that the sampling effort should be focused on data for living taxa.
\end{enumerate}

@@

\item{\textcolor{blue}{\textbf{Introduction:} Needs stronger emphasis on what specifically is unique about this study compared to what the many other studies on missing data have done and found.
Clearly outlining the questions in this study that are left open from previous research is important because some of the findings in this study suggest that there may be more error associated with missing data than previous research has suggested.}}

% Needs stronger emphasis on what specifically is unique about this study compared to what the many other studies on missing data have done and found.
% To do (see reviewer 1 specific comment 13) - links to major comment one on novelty

% Clearly outlining the questions in this study that are left open from previous research is important because some of the findings in this study suggest that there may be more error associated with missing data than previous research has suggested.

@@

\item{\textcolor{blue}{\textbf{Abbreviation:} One suggestion is to consider abbreviating Maximum Likelihood and Bayesian Inference as ML and BI respectively early in the methods and use throughout.}}

We do understand the reviewers concern are agree that ML and BI are two commonly used abbreviations.
Nonetheless, we would rather prefer not abbreviate these terms: in fact, even though they might add some concision to the sentences where Maximum Likelihood and/or Bayesian Inference are stated, we think they might add confusion or difficulties to the whole paper.
Spelling out the abbreviation might be locally less concise (i.e. in the sentence) but at least are completely clear and does not ask the reader to look back for the actual meaning of such or such abbreviation (Berlin 2013; Radiology; \url{http://pubs.rsna.org/doi/abs/10.1148/radiol.12121776}).
However, we leave this decision to the editor if she judges it better to abbreviate those terms. % TG: or maybe this statement is useless? Of course we leave it to the editor...

@@

\item{\textcolor{blue}{\textbf{Methodology - True tree:} I suggest it warrants comparing at least the 'best' tree to the 'true' tree.}}

We compared both the ``best'' tree and the ``missing-data'' trees to the ``true'' tree as suggested by both reviewer.
We did not, however, include the results to the main text (available in appendix B; figure B1 and appendix C; figures C6 and C7).
We do agree with the point raised with both reviewers on comparing the true tree to the best/missing-data trees.
As reviewer 1 states: "[this analysis can give] insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it.".
However, even though we agree with the comments of both reviewer that knowing the difference between the true and best/missing-data trees is important, we do not think that it "adds an additional layer of complexity that can't be accounted for via simulation" (reviewer 2, comment 1).
In this simulations, we had to make a choice between (1) simulating larger matrices but with less chains (i.e. less variance) or (2) smaller matrices with more variance.
We deliberately made the second choice because our main question was "what is the effect of our parameters on recovering the topology".
In fact, regarding our problematic, we maintain that comparing the best tree to the missing-data trees is the best approach to answer this specific question for two reasons:
\begin{enumerate}
\item{The true tree is always unknown:} as we state in the manuscript: "In real world situations, the ``true'' tree is never available" (line 242 in the first manuscript and @@@ in the revision).
\item{The true tree is just the starting point of the simulations:} in fact, we use the true tree only as a starting point for our simulations. Once generated, the true tree is fed to a cascade of algorithms and softwares (i.e. seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes) described in section 2.1 and 2.3 in order to get the best tree and the missing-data trees. Comparing the latests to the true tree will answer to the problematic "what is the effect of our missing-data parameters and the cascade of algorithms and softwares on recovering the true topology?".
\end{enumerate}
Therefore we added the second justification to the main text (see below this response) and added these analysis (true vs. best/missing-data trees) to the supplementary for potential readers that might be interested by this specific problematic.

@@

We added the following sentence to the main text to clarify and link to the supplementary analysis:
"In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices (i.e. diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." (line @@@).

\item{\textcolor{blue}{\textbf{Methodology - Starting tree:} The authors should at least include some analysis in which no starting tree is given, the max. likelihood 'best' tree is given, and / or the 'best' tree is given but with some perturbation.
I fear that one of the main results may be biased due to this one methodological detail - a strong prior on tree topology.
The justification given was 'to speed up the Bayesian analysis' which does not seem adequate, given these datasets are small.}}

We dealt with this legitimate suggestion an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. 
We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree.
We found no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
We made these data available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a precision in the manuscript:

@@

"We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
Note that a starting is not a Bayesian prior on topology \textit{per se} and did not significantly affected topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line @@@

\item{\textcolor{blue}{\textbf{Results - Bhattacharyya meaning:} Some descriptive statistics are given, but it would help to tell us if these differences are meaningful.
For example, on line 414 the Bhattacharrya Coefficients are given suggesting RF distances are lower for ML than BI trees at 0.69, 0.48, and 0.66 - but if this coefficient represents the probability of overlap between distributions then shouldn't the coefficients be $<$0.05 to say that one is lower than the other?
In the table, mean and median values of $<$0.01 and $>$0.85 are highlighted as being the most extreme.
Later the focus on Bhattacharyya Coefficients is 1 when distributions overlap completely and 0 when they do not - should this be the criterion?
Just make it clear early on so the readers knows what to think when they see these numbers. }}

We clarified the interpretation of the Bhattacharyya Coefficient throughout the manuscript.
We advocate that two distributions are significantly different when their probability of overlapping (the Bhattacharyya Coefficient) is $>$0.05 and that they are significantly similar when this probability is $>$0.95.
Even though other values (e.g. 0.85) are not conclusive, we still reported them to show the relative overlap of the distributions (i.e. two distributions with a coefficient of 0.85 are not significantly similar but are closer to each other than two distributions with a coefficient of 0.15; which are themselves not significantly different but still less closer to each other than the previous ones).
We added the following statement to the main text and to the appendix B:

"When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
When this coefficient is $>$0.95 both distributions are significantly similar.
Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions." line (@@@)

@@

\item{\textcolor{blue}{\textbf{Results - Bayesian consensus vs. Bayesian posterior:} One major finding that was not fully addressed was that the distances between the BI posterior distributions of trees do not overlap the BI best or consensus trees.
It seems from the tables in Appendix C that the topology distance metrics for Bayesian posterior distribution are similar to those for the Bayesian consensus, albeit usually lower for RF distance than the Tr distance.
In Figure 4, it seems like the distributions of metrics based on CIs for Bayesian consensus and posterior overlap up until >25\% missing data.
How then in comparisons of the distributions are they always so low (Table 1 max prob. of overlapping distributions = 0.11)?}}

The difference between the Bayesian consensus and the Bayesian posterior trees can be due to two factors:
\begin{enumerate}
\item first, the way we assess topological recovery is slightly different in each case: for the Bayesian consensus trees, we simply compared the "missing-data" consensus tree to the "best" consensus tree; for the Bayesian posterior trees however, we randomly compared 1000 "missing-data" Bayesian posterior trees to 1000 "best" Bayesian posterior trees.
Therefore, even if the comparison methodology is the same, the fact that the second methods draws random trees, we expect the topological recovery score to be "blurred" in comparison to other comparisons that do not involve random comparisons.
\item secondly, the Bayesian posterior trees are always resolved and thus, in the same way as the Maximum Likelihood tree, more likely to contain "false-positive" nodes. On the other hand, because the consensus tree is a strict majority consensus rule, only nodes with a a good support will be represented in the tree and all other nodes will be collapsed. This reduces the amount of "false-positive" nodes (since an unresolved tree is always better than a wrongly resolved one) and increases clade conservation but also increases the placement of wild-card taxa (since it is likely that some Triplets will just not exist in a poorly resolved consensus tree).
\end{enumerate}
Both points were discussed in the former manuscript lines 566 and 573 (or lines @@@ and @@@ in the revised version).
We expanded the paragraph to:

"Additionally, the Bayesian posterior trees performed way worth than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
This can be due to the fact that the Bayesian posterior trees are always resolved and thus more likely to contain wrongly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
Conversely, the Bayesian consensus tree might not resolve nodes that are badly supported and thus more likely to contain only correctly resolved nodes (i.e. increasing the Normalised Robinson-Foulds metric)." lines (@@@)

Concerning the last question of this comment: "How then in comparisons of the distributions are they always so low (Table 1 max prob. of overlapping distributions = 0.11)?" when looking at all parameters separately (figure 4), there is no clear overlap between the Bayesian consensus trees and the Bayesian posterior trees (median Bhattacharyya Coefficient of 0.01; table 1).
The maximum observed Bhattacharyya Coefficient (0.11; table 1) is probably due to the combination of parameters: when all the parameters have many missing data, the probability of overlap between trees was slightly higher (0.11) but still really low.
For the Normalised Triplets distance however, the overlap is higher (median Bhattacharyya Coefficient of 0.59; table 1) as illustrated in figure 4 (i.e. much more overlap between the distributions).

@@

% TG: I'm confused for the second part of the question:
% TG: In figure 4 and in Table 1, if you look at the Bayesian consensus (grey) trees and the Bayesian posterior trees (orange) for RF, they do not really overlap. Ever. The 0.11 probability must come from the parameters combinations that, when is really shitty must show a little bit of overlap. But for each parameter one by one: nope. No overlap.
% TG: On the other side, in figure 4 and Table 1, for Bayesian consensus and Bayesian posterior for Tr, the do overlap much more. And that's what's in the Table 1 (min 0.21; median 0.56; and max 0.87).
% TG: Did the reviewer just miss read the labels of the Table 1 or am I miles away from what he means?

\item{\textcolor{blue}{\textbf{Results - Bayesian Triplets:} Another important result that is not mentioned is that using the wildcard taxon metric (Tr), Bayesian consensus trees do quite poorly with missing data.
Values in Figures 4 and 5 clearly drop below zero, although negative values are not illustrated on the axis, and this indicates that the tree topologies are no better than, or even worse than, random trees.
It is especially true when fossils are missing data, and that is going to be an important finding with practical application but must be addressed. }}

We modified the figures 4 and 5 to show Normalised Triplets metric values below 0 and added a dashed line at 0 to underline the difference in scale between both metrics.
As mentioned above (reviewer 1 comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
This has two related effects.
The first one being to retain "good" clade conservation (i.e. avoiding "false-positive" nodes by collapsing unsupported ones).
And the second one is that when these clades are collapsed, it reduces the amount of available triplets and thus increases the "wild-card taxa" score (Normalised Triplets metric).
In fact, when a clade ((a;b);c); is collapsed to (a;b;c), the whole clade is conserved (a, b and c are still in a common clade) however they also all become wild-card taxa: the position of a regarding b and c is now incorrect.

We added a precision on this point in the...
% TG: add a paragraph in the results about the wild-card fossils

% TG: STOPPED HERE.

@@

\item{\textcolor{blue}{\textbf{line 231-235:} streamline, e.g., 'The minimal dataset included in this study was 5\% of the morphological matrix for any taxon'.}}

We modified the sentence as follows: "To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon." (line @@@).

\item{\textcolor{blue}{\textbf{line 237-246:} I understand the rational presented by the authors for comparing missing data trees to the best tree inferred from max. likelihood/Bayesian techniques rather than compare missing data trees to the true tree used to simulate the data. I still feel it would be an important contribution to this paper, however, to compare the missing data trees to the true tree to evaluate the effects of missing data on the tree inference. Also it may lend insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it. }}

We dealt with this comment above (Reviewer 1, major comment 6) and below (Reviewer 2 comment 1).

@@

\item{\textcolor{blue}{\textbf{line 273:} 'Bayesian Inference'}}

We changed the title as suggested.

### 2015-07-31 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
updated response

\subsubsection{Robinson-Foulds distance}
The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and \ul{twice the number} of shared clades across two trees.
The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} (see Appendix B for calculation details).
This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
We normalised this metric following Bogdanowicz's Normalised Tree Similarity (NTS) method \citep{Bogdanowicz2012}.
This method scales any tree comparison metric using the mean distance between 1000 random trees (see Appendix B for the calculation details).
This method is a generalisation of the topological accuracy method \citep{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.
In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
Note that once rescaled, the Normalised Robinson-Foulds metric is a measure of similarity, rather than of distance like the original Robinson-Foulds metric.

@@

We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
The Bhattacharyya Coefficient is the probability of overlap between two distributions \citep{Bhattacharyya} (see Appendix B for calculation details).
Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
When this coefficient is $>$0.95 both distributions are significantly similar.
Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions.
\hl{To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $N_{C}$ = 100 etc. (see Figure 2 for more details).}
This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
We performed this procedure separately for each tree inference and uncertainty method.
When two combinations of missing data parameters have a similar ability to recover the ``best'' tree topology the Bhattacharyya Coefficient will be close to one.
Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
\hl{Because of the difficulties to represent so many pairwise comparisons in a meaningful way, we summarized as a heat map of Bhattacharyya Coefficients (see Figure 6).
In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).}

@@

The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds metric) but greater on the displacement of wildcard taxa (Normalised Triplets metric; Figure 4 and Figure 5).
This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
When the fossil taxa have less data (e.g. $M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
Therefore a majority consensus position is unlikely to exist (i.e. every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
\ul{In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.}
Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxon and decrease the Normalised Triplets metric.
Therefore, the $M_{F}$ parameter is likely to affect the Normalised Robinson-Foulds metric less than the Normalised Triplets metric for the Bayesian consensus trees.
Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).
In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level.
We argue that using the Bayesian consensus tree topology is preferable because it is more conservative \citep[e.g.][]{pattinsonphylogeny2014}.

\subsubsection{Number of morphological characters ($N_{C}$)}
Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the ``best'' tree topology.
We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
Conversely, removing 10\% of morphological characters ($N_{C}$) genuinely removes 10\% of data in the matrix.
Nonetheless, the effect of removing characters on the ability to recover the ``best'' tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.

Additionally, the number of morphological characters determines the size the of matrix.
This can affect our ability to recover the ``best'' tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \hl{(}\citealt{bremer1992phylogeny,patterson1993congruence}\hl{; see }\citealt{masters2002lack}\hl{ for a empirical example)}.
\ul{The sizes of our data matrices were constrained by the performance of our protocol}: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
Although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
Therefore, our simulations reflect realistic parameters.
Nonetheless, the use of probabilistic methods (i.e. Maximum Likelihood or Bayesian) and the M\textit{kv} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.

@@

It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best'' tree topology than Maximum Likelihood and Bayesian consensus trees.
This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $N_{C}$ = 0\%; Figure 4).
This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
This will obviously add noise to the results for these methods.

>>>>

\subsubsection{Robinson-Foulds distance}
The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and \hl{twice the number} of shared clades across two trees.
The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} (see Appendix B for calculation details).
This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
We normalised this metric following Bogdanowicz's Normalised Tree Similarity (NTS) method \citep{Bogdanowicz2012}.
This method scales any tree comparison metric using the mean distance between 1000 random trees (see Appendix B for the calculation details).
This method is a generalisation of the topological accuracy method \citep{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.
In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
Note that once rescaled, the Normalised Robinson-Foulds metric is a measure of similarity, rather than of distance like the original Robinson-Foulds metric.

@@

We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
The Bhattacharyya Coefficient is the probability of overlap between two distributions \citep{Bhattacharyya} (see Appendix B for calculation details).
Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
\hl{When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
When this coefficient is $>$0.95 both distributions are significantly similar.
Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions.}
\hl{To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $N_{C}$ = 100 etc. (see Figure 2 for more details).}
This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
We performed this procedure separately for each tree inference and uncertainty method.
When two combinations of missing data parameters have a similar ability to recover the ``best'' tree topology the Bhattacharyya Coefficient will be close to one.
Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
\hl{Because of the difficulties to represent so many pairwise comparisons in a meaningful way, we summarized as a heat map of Bhattacharyya Coefficients (see Figure 6).
In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).}

@@

The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds metric) but greater on the displacement of wildcard taxa (Normalised Triplets metric; Figure 4 and Figure 5).
This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
When the fossil taxa have less data (e.g. $M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
Therefore a majority consensus position is unlikely to exist (i.e. every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
\hl{In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.}
Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxon and decrease the Normalised Triplets metric.
Therefore, the $M_{F}$ parameter is likely to affect the Normalised Robinson-Foulds metric less than the Normalised Triplets metric for the Bayesian consensus trees.
Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).
In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level.
We argue that using the Bayesian consensus tree topology is preferable because it is more conservative \citep[e.g.][]{pattinsonphylogeny2014}.

\subsubsection{Number of morphological characters ($N_{C}$)}
Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the ``best'' tree topology.
We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
Conversely, removing 10\% of morphological characters (\hl{i.e. $N_{C}$ = 90}) genuinely removes 10\% of data in the matrix.
Nonetheless, the effect of removing characters on the ability to recover the ``best'' tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.

Additionally, the number of morphological characters determines the size the of matrix.
This can affect our ability to recover the ``best'' tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \hl{(}\citealt{bremer1992phylogeny,patterson1993congruence}\hl{; see }\citealt{masters2002lack}\hl{ for a empirical example)}.
\hl{The sizes of our data matrices were constrained by the performance of our protocol}: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
Although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
Therefore, our simulations reflect realistic parameters.
Nonetheless, the use of probabilistic methods (i.e. Maximum Likelihood or Bayesian) and the M\textit{kv} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.

@@

The Bayesian consensus trees, however, performs badly for the Normalised Triplets metric: some parameters combinations, especially when the $M_F$ parameter reaches 75\% missing data, leads to negative values (figure 5).
Normalised Triplets metric values below 0 means that the placement of some taxa is worth than expected by just randomly placing this taxa in the tree.
This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees [CONTINUE LATER ON].

It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best'' tree topology than Maximum Likelihood and Bayesian consensus trees.
This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $N_{C}$ = 100; Figure 4).
This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
\hl{Additionally, the Bayesian posterior trees performed way worth than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
This can be due to the fact that the Bayesian posterior trees are always resolved and thus more likely to contain wrongly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
Conversely, the Bayesian consensus tree might not resolve nodes that are badly supported and thus more likely to contain only correctly resolved nodes (i.e. increasing the Normalised Robinson-Foulds metric).}
